---
title: "HW2"
author: "Tara Jagadeesh"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
data<- read_csv("HW_#2.csv")

```

Question 1.

a) Make a plot for the demand curves
```{r}
demand_1<- ggplot(data, aes(x = Q_Low_Gallons_per_Day, y = Price_Dollars))+
  geom_point() +
  stat_smooth(method = "lm")
demand_1

demand_2<- ggplot(data, aes(x = Q_High_Gallons_per_Day, y = Price_Dollars))+
  geom_point() +
  stat_smooth(method = "lm")
demand_2
```

b) Linear model of the demand curves
```{r}
demand_low_lm<- lm(data$Price_Dollars ~ data$Q_Low_Gallons_per_Day)
plot(demand_low_lm)
summary(demand_low_lm)

demand_high_lm<- lm(data$Price_Dollars ~ data$Q_High_Gallons_per_Day)
plot(demand_high_lm)
summary(demand_high_lm)

# Low demand coefficients
#P = -0.0001355174(Q) + 21.99085
alpha_low <- demand_low_lm[["coefficients"]][["(Intercept)"]]
beta_low <- demand_low_lm[["coefficients"]][["data$Q_Low_Gallons_per_Day"]]
# p = beta(q) + alpha

# High demand coefficients
#P = -0.0001296638(Q) + 23.39144
alpha_high <- demand_high_lm[["coefficients"]][["(Intercept)"]]
beta_high <- demand_high_lm[["coefficients"]][["data$Q_High_Gallons_per_Day"]]
# p = beta(q) + alpha

#Equation for low demand group
Q_low1<- (1/beta_low)
Q_low2<- (alpha_low/-beta_low)
# Q_low = Q_low1*P + Q_low2


#Equation for high demand group
Q_high1<- (1/beta_high)
Q_high2<- (alpha_high/-beta_high)
# Q_high = Q_high1*P + Q_high2

Q_agg1<- Q_low1+Q_high1 # aggreagte low demand
Q_agg2<- Q_low2+Q_high2 # aggregate high demand

P_agg1<- 1/Q_agg1 # price for low demand
P_agg2<- -Q_agg2/Q_agg1 # price for high demand

# Supply curve
supply<- 5/((5-P_agg2)/P_agg1)
supply
# equation of supply  P = 0.00001871
```


c) Demand curve functions
```{r}
#low demand curve
demand_low<- function(x){alpha_low+beta_low*x}

# high demand curve
demand_high<- function(x){alpha_high+beta_high*x}

# aggregate demand curve
demand_agg<- function(x){P_agg2+P_agg1*x}

# Supply curve
supply_curve<- function(x){supply*x}

# marginal external cost curve
mec<- function(x){2*x}

# agregate equation
agg_intercept<- (alpha_high/beta_high)+(alpha_low/beta_low)
agg_slope<- abs((-1/beta_high) + (-1/beta_low))
```

d) Demand curve graphs
```{r}
# optimum point on th graph

opt<- data.frame(x = c(0:267176.2), y = c(5))
x<- 70000:500000

graph_1<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_curve, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))

graph_1


```

e) Aggregate consumer benefit
```{r}
#Equilibrium pt: demand_agg = supply_curve
#P = 5
#Q = 267217.062558

#Y intercept of agg demand
demand_agg(0) #22.70661

#Create a function for agg demand at the equilibrium point
demand_agg2<- function(x){P_agg2+P_agg1*x - 5}

#Find the area under the curve via integrating
agg_cs <- integrate(demand_agg2, 0, 267217.062558)
agg_cs #2365754

#Double check the area using shapes
agg_height <- 22.70661 - 5
agg_base <- 267217.062558
agg_cs <- (1/2)*agg_base*agg_height #2365754.1560 WOOOOOO THEYRE EQUAL 

#Aggregate consumer surplus  = 2365754.1560


##wait do we have to account for the kink??? 
```

f) Producer surplus
```{r}
# Equilibrium demand_agg = supply_curve
# P = 5
# Q = 267217.062558

supply_curve(0) # = 0

# geometric PS

p_agg_hight<- 5
p_agg_base<- 267217.062558
p_agg_ps<- (1/2)*p_agg_hight*p_agg_base
p_agg_ps

# aggregate PS = 668042.7

```

g) Environmental cost
```{r}
# mec = 2

tec_base<- 267217.062558
tec_height<- 2
tec<- tec_base*tec_height
tec
# total environmental cost = 534434.1
```


Question 2: Distribution of benefits

a) Low demand consumer benefit at P = 5
```{r}
#5 = -0.0001355174(Q) + 21.99085
#Q = 125377.626784

demand_low(125377.626784)

#Create a function for low demand at the current price
demand_low2<- function(x){alpha_low+beta_low*x - 5}

#Find the area under the curve via integrating
demandint_low <- integrate(demand_low2, 0, 125377.626784)
demandint_low #1065137

#Double check the area using shapes
low_height <- 21.99085 - 5
low_base <- 125377.626784
low_cs <- (1/2)*low_base*low_height #1065136 

#Low demand benefit = 1065136

```

b) High demand consumer benefit at P = 5
```{r}
#5 = -0.0001296638(Q) + 23.39144
#Q = 141839.433982

demand_high(141839.433982)

#Create a function for low demand at the current price
demand_high2<- function(x){alpha_high+beta_high*x - 5}

#Find the area under the curve via integrating
demandint_high <- integrate(demand_high2, 0, 141839.433982)
demandint_high #1304316

#Double check the area using shapes
high_height <- 23.39144 - 5
high_base <- 141839.433982
high_cs <- (1/2)*high_base*high_height #1304315.7199

#High demand benefit = 1304315.7199

```


Question 3: Gas tax

a) New quantity of gas
```{r}
# make a new supply cnurve with the 0.5 tax
supply_curve_tax<- function(x){supply*x+0.5}

# graph the new supply to see what it looks like

graph_tax<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_curve_tax, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))

graph_tax

# looks correct (supply shifted up by 0.5)

# find the new equilibrium with the new supply curve

q_opt<- uniroot(function(x) demand_agg(x) - supply_curve_tax(x), range(x))$root
q_opt

# New equilibrium qunatity = 261333

p_opt<- supply_curve_tax(q_opt)
p_opt

# new equilibrium price = 5.3899

# Q_tax = 261333
# P_tax = 5.3899


```

b) New gas price
```{r}
p_opt<- supply_curve_tax(q_opt)
p_opt

# new equilibrium price = 5.3899
```


c) New high consumer benefit
```{r}
#5.3899 = -0.0001296638(Q) + 23.39144
# Q = 138836.495

demand_high3<- function(x){alpha_high+beta_high*x - 5.3899}

demandint_high_tax <- integrate(demand_high3, 0, 138836.495)
demandint_high_tax

# geometric

demand_high(0)
# y - int = 23.39144

demand_high_base_tax<- 23.39144 - 5.3899
demand_high_height_tax<- 138836.495
demand_high_cs_tax<- (1/2)*demand_high_height_tax*demand_high_base_tax
demand_high_cs_tax

# CS for high demand with the new tax = 1249635

```

d) New low consumer benefit
```{r}
#5.3899 = -0.0001355174(Q) + 21.99085
# Q = 119,179.846

# geometric 

demand_low(0)
# y - int = 21.99085

demand_low_base_tax<- 21.99085 - 5.3899
demand_low_height_tax<- 119179.846
demand_low_cs_tax<- (1/2)*demand_low_base_tax*demand_low_height_tax
demand_low_cs_tax

# CS for low demand users with the new tax = 989249.3

```

e) New producer benefit
```{r}
# geometric 

supply_curve_tax(0)
# y - int = 0.5

supply_tax_base<- 5.3899-0.5
supply_tax_height<- 261333
ps_tax<- (1/2)*supply_tax_base*supply_tax_height
ps_tax

# PS under the new tax = 638946.1

```

f) Total envr damage
```{r}

mec_base_tax<- 261333
mec_height_tax<- 2
tec_tax<- mec_height_tax*mec_base_tax
tec_tax

# Total environmental cost under the new tax = 522666

```

g) Total tax revenue 
```{r}
# the length of the tax revenue rectanlge will just be the optinaml Q with the tax
# the height of the taxx revenue will be the difference between the optimal price before and after the tax
tax_rev_base<- 261333
tax_rev_height<- 5.3899 - 5
tax_rev<- tax_rev_height*tax_rev_base
tax_rev

# the revenue from the tax =  101893.7

```

4. Now, assume that all revenue from a tax will be used for infrastructure repairs,
and that the benefit of this is proportional to the amount you drive (i.e. to the gas
you consume). Also assume that “Low” income consumers bear all
environmental costs. For a range of gas taxes (ranging from $0 - $5.00/gal),
calculate the effects of the tax on:


```{r}
# set up the problem

aggfxn_org <- function(Q){-0.00006626299*Q + 22.70661}
suppfxn_org <- function(Q){0.00001871*Q}

taxfxn_0.5 <- function(Q){0.00001871*Q + 0.5}
taxfxn_1 <-  function(Q){0.00001871*Q + 1}
taxfxn_2.5 <- function(Q){0.00001871*Q + 2.5}
taxfxn_5 <- function(Q){0.00001871*Q + 5}

# Quantity at each tax level
eq_org <- uniroot(function(x) aggfxn_org(x) - suppfxn_org(x), range(x))$root
eq_org
#  q =  267221.5
eq_0.5 <- uniroot(function(x) aggfxn_org(x) - taxfxn_0.5(x), range(x))$root
eq_0.5
# q = 261337.3
eq_1 <- uniroot(function(x) aggfxn_org(x) - taxfxn_1(x), range(x))$root
eq_1
# q = 255453.1
eq_2.5 <- uniroot(function(x) aggfxn_org(x) - taxfxn_2.5(x), range(x))$root
eq_2.5
# q = 237800.4
eq_5 <- uniroot(function(x) aggfxn_org(x) - taxfxn_5(x), range(x))$root
eq_5
# q = 208379.3


# Market price at each tax level
aggfxn_org(eq_org) # p = 4.999714
taxfxn_0.5(eq_0.5) # p = 5.38962

taxfxn_1(eq_1) # p = 5.779527
taxfxn_2.5(eq_2.5) # p = 6.949245
taxfxn_5(eq_5) # p = 8.898776




#Low consumer surplus
low_eq_1<- uniroot(function(x) demand_low(x) - 5.779527 , range(x))$root
low_eq_1
# q = 119625.4
low_eq_2.5<- uniroot(function(x) demand_low(x) - 6.949245 , range(x))$root
low_eq_2.5
# q = 110993.9
low_eq_5<- uniroot(function(x) demand_low(x) - 8.898776 , range(x))$root
low_eq_5
# q = 96608.09



graph_tax_dif<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_curve, color = "pink")+
  stat_function(fun = taxfxn_1, color = "purple")+
  stat_function(fun = taxfxn_2.5, color = "orange")+
  stat_function(fun = taxfxn_5, color = "yellow")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))

graph_tax_dif







```

B. High demand consumer surplus
```{r}
# Overall welfare of high demand = high demand consumer surplus

#Tax of $1
high_eq_1<- uniroot(function(x) demand_high(x) - 5.779527 , range(x))$root
high_eq_1 # q = 135827.6

high_base_1 <- 135827.6
high_height_1 <- demand_high(0) - demand_high(135827.6)
high_CS_1 <- (1/2)*high_base_1*high_height_1
high_CS_1 #High demand CS = 1196092

#Tax of $2.5
high_eq_2.5<- uniroot(function(x) demand_high(x) - 6.949245 , range(x))$root
high_eq_2.5 # q = 126806.4

high_base_2.5 <- 126806.4
high_height_2.5 <- demand_high(0) - demand_high(126806.4)
high_CS_2.5 <- (1/2)*high_base_2.5*high_height_2.5
high_CS_2.5 #High demand CS = 1042488

#Tax of $5
high_eq_5<- uniroot(function(x) demand_low(x) - 8.898776 , range(x))$root
high_eq_5 # q = 96608.09

high_base_5 <- 96608.09
high_height_5 <- demand_high(0) - demand_high(96608.09)
high_CS_5 <- (1/2)*high_base_5*high_height_5
high_CS_5 #High demand CS = 809930.5

# HIGH DEMAND CONSUMER SUMMARY:
#(No tax  ): High demand welfare is $1304315
#($0.5 Tax): High demand welfare is $1249635
#($1 Tax  ): High demand welfare is $1196092
#($2.5 Tax): High demand welfare is $1042488
#($5 Tax  ): High demand welfare is  $809930

```

B. Overall wellfare of low demand consumers
```{r}
# Overall welfare of low demand = low demand consumer surplus - environmental costs

#### ALERT: I just copy/pasted for now! Still have to change values below


#Tax of $1
env_cost_1<- mec(119625.4)
env_cost_1

low_eq_1<- uniroot(function(x) demand_low(x) - 5.779527 , range(x))$root
low_eq_1 # q = 119625.4

low_base_1 <- 119625.4
low_height_1 <- demand_low(0) - demand_low(119625.4)
low_CS_1 <- ((1/2)*low_base_1*low_height_1)-env_cost_1
low_CS_1 #Low demand CS = 730392.3

#Tax of $2.5
env_cost_2.5<- mec(110993.9)
low_eq_2.5<- uniroot(function(x) demand_low(x) - 6.949245 , range(x))$root
low_eq_2.5 # q = 110993.9

low_base_2.5 <- 110993.9
low_height_2.5 <- demand_low(0) - demand_low(110993.9)
low_CS_2.5 <- ((1/2)*low_base_2.5*low_height_2.5)-env_cost_2.5
low_CS_2.5 #Low demand CS = 612775.4

#Tax of $5
env_cost_5<- mec(96608.09)
low_eq_5<- uniroot(function(x) demand_low(x) - 8.898776 , range(x))$root
low_eq_5 # q = 96608.09

low_base_5 <- 96608.09
low_height_5 <- demand_low(0) - demand_low(96608.09)
low_CS_5 <- ((1/2)*low_base_5*low_height_5)-env_cost_5
low_CS_5 #High demand CS = 439184.1

# LOW DEMAND CONSUMER SUMMARY:
#(No tax  ): Low demand welfare is $1065136 
#($0.5 Tax): Low demand welfare is $989249.3
#($1 Tax  ): Low demand welfare is $730392.3
#($2.5 Tax): Low demand welfare is $612775.4
#($5 Tax  ): Low demand welfare is $439184.1

```

C.  Producer surplus for gas producers
```{r}
# make new equilibrium points for the three taxes with the agg demand

# Tax of $1
supply_curve(255453)
# p = 4.779877
ps_eq_1<- uniroot(function(x) demand_agg(x) - 5.779527, range(x))$root
ps_1
# q = 255453

ps_1_base<- 255453
ps_1_height<- 4.779877
ps_1<- (1/2)*ps_1_base*ps_1_height
ps_1
# ps for a 1$ tax = 610517


# Tax  for $2.50
supply_curve(237800.3)
# p = 4.449571
ps_eq_2.5<- uniroot(function(x) demand_agg(x) - 6.949245, range(x))$root
# q = 237800.3

ps_2.5_base<- 237800.3
ps_2.5_height<- 4.449571
ps_2.5<- (1/2)*ps_2.5_base*ps_2.5_height
ps_2.5
# PS under a $2.5 tax = 529054.7

# Tax for $5
supply_curve(208379.2)
# p = 3.899062
ps_eq_5<- uniroot(function(x) demand_agg(x) -8.898776, range(x))$root
ps_eq_5
# q =  208379.2

ps_5_base<- 208379.2
ps_5_height<- 3.899062
ps_5<- (1/2)*ps_5_base*ps_5_height
ps_5
# PS under the $5 tax = 406241.7


# PRODUCER SURPLUS SUMMARY:
# No Tax: 668042.7
# $0.5 Tax: 638946.1
# $1 Tax: 610517
# $2.5 Tax: 529054.7
# $5 Tax: 406241.7


```


5. Lowered demand curves
```{r}
#New low demand curve
lowdemand_new <- function(Q){-0.0001355174*Q + 21.99085/2}
#New low demand curve
highdemand_new <- function(Q){-0.0001296638*Q + 23.39144/2}
#New agg demand curve
aggdemand_new <- function(Q){-0.00006626299*Q + 22.70661/2}

new_graph <- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = lowdemand_new, color = "darkblue")+
  stat_function(fun = highdemand_new, color = "darkorange")+
  stat_function(fun = aggdemand_new, color = "darkgreen")+
  stat_function(fun = supply_curve, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))
new_graph

# new optimal points

opt_agg_new<- uniroot(function(x) aggdemand_new(x) - suppfxn_org(x), range(x))$root
opt_agg_new
# q = 133610.8

opt_price<- aggdemand_new(133610.8)
opt_price
# p = 2.499854

# new agg optimum point
# p = 2.499854
# q = 133610.8

# quantity consumed by the low demand
lowdemand_new(2.499854) # q = 10.99509

# quantnity consumed by the high demand
highdemand_new(2.499854) # q = 11.6954

#environmental damage 
new_env_cost<- mec(133610.8)
new_env_cost # 267221.6

```

Gas consumption by “High” income consumers
b. Gas consumption by “Low” income consumers
c. Gas price
d. Environmental damage from gasoline





