---
title: "HW#4"
author: "Tyler Cobian"
date: "4/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
data<- read_csv("HW_#2.csv")

```

Question 1.

a) Make a plot for the demand curves
```{r}
demand_1<- ggplot(data, aes(x = Q_Low_Gallons_per_Day, y = Price_Dollars))+
  geom_point() +
  stat_smooth(method = "lm")
demand_1

demand_2<- ggplot(data, aes(x = Q_High_Gallons_per_Day, y = Price_Dollars))+
  geom_point() +
  stat_smooth(method = "lm")
demand_2
```

b) Linear model of the demand curves
```{r}
demand_low_lm<- lm(data$Price_Dollars ~ data$Q_Low_Gallons_per_Day)
plot(demand_low_lm)
summary(demand_low_lm)

demand_high_lm<- lm(data$Price_Dollars ~ data$Q_High_Gallons_per_Day)
plot(demand_high_lm)
summary(demand_high_lm)

# Low demand coefficients
#P = -0.0001355174(Q) + 21.99085
alpha_low <- demand_low_lm[["coefficients"]][["(Intercept)"]]
beta_low <- demand_low_lm[["coefficients"]][["data$Q_Low_Gallons_per_Day"]]
# p = beta(q) + alpha

# High demand coefficients
#P = -0.0001296638(Q) + 23.39144
alpha_high <- demand_high_lm[["coefficients"]][["(Intercept)"]]
beta_high <- demand_high_lm[["coefficients"]][["data$Q_High_Gallons_per_Day"]]
# p = beta(q) + alpha

#Equation for low demand group
Q_low1<- (1/beta_low)
Q_low2<- (alpha_low/-beta_low)
# Q_low = Q_low1*P + Q_low2


#Equation for high demand group
Q_high1<- (1/beta_high)
Q_high2<- (alpha_high/-beta_high)
# Q_high = Q_high1*P + Q_high2

Q_agg1<- Q_low1+Q_high1 # aggreagte low demand
Q_agg2<- Q_low2+Q_high2 # aggregate high demand

P_agg1<- 1/Q_agg1 # price for low demand
P_agg2<- -Q_agg2/Q_agg1 # price for high demand

# Supply curve
supply<- 5/((5-P_agg2)/P_agg1)
supply
# equation of supply  P = 0.00001871
```


c) Demand curve functions
```{r}
#low demand curve
demand_low<- function(x){alpha_low+beta_low*x}

# high demand curve
demand_high<- function(x){alpha_high+beta_high*x}

# aggregate demand curve
demand_agg<- function(x){P_agg2+P_agg1*x}

# Supply curve
supply_curve<- function(x){supply*x}

# marginal external cost curve
mec<- function(x){2 + 0*x}

# agregate equation
agg_intercept<- (alpha_high/beta_high)+(alpha_low/beta_low)
agg_slope<- abs((-1/beta_high) + (-1/beta_low))
```

d) Demand curve graphs
```{r}
# optimum point on th graph

opt<- data.frame(x = c(0:267176.2), y = c(5))
x<- 70000:500000

graph_1<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_curve, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))

graph_1


```

e) Aggregate consumer benefit
```{r}
#Equilibrium pt: demand_agg = supply_curve
#P = 5
#Q = 267217.062558

#Y intercept of agg demand
demand_agg(0) #22.70661

#Create a function for agg demand at the equilibrium point
demand_agg2<- function(x){P_agg2+P_agg1*x - 5}

#Find the area under the curve via integrating
agg_cs <- integrate(demand_agg2, 0, 267217.062558)
agg_cs #2365754

#Double check the area using shapes
agg_height <- 22.70661 - 5
agg_base <- 267217.062558
agg_cs <- (1/2)*agg_base*agg_height #2365754.1560 WOOOOOO THEYRE EQUAL 

#Aggregate consumer surplus  = 2365754.1560


##wait do we have to account for the kink??? 
```

f) Producer surplus
```{r}
# Equilibrium demand_agg = supply_curve
# P = 5
# Q = 267217.062558

supply_curve(0) # = 0

# geometric PS

p_agg_hight<- 5
p_agg_base<- 267217.062558
p_agg_ps<- (1/2)*p_agg_hight*p_agg_base
p_agg_ps

# aggregate PS = 668042.7

```

g) Environmental cost
```{r}
# mec = 2

tec_base<- 267217.062558
tec_height<- 2
tec<- tec_base*tec_height
tec
# total environmental cost = 534434.1
```


Question 2: Distribution of benefits

a) Low demand consumer benefit at P = 5
```{r}
#5 = -0.0001355174(Q) + 21.99085
#Q = 125377.626784

demand_low(125377.626784)

#Create a function for low demand at the current price
demand_low2<- function(x){alpha_low+beta_low*x - 5}

#Find the area under the curve via integrating
demandint_low <- integrate(demand_low2, 0, 125377.626784)
demandint_low #1065137

#Double check the area using shapes
low_height <- 21.99085 - 5
low_base <- 125377.626784
low_cs <- (1/2)*low_base*low_height #1065136 

#Low demand benefit = 1065136

```

b) High demand consumer benefit at P = 5
```{r}
#5 = -0.0001296638(Q) + 23.39144
#Q = 141839.433982

demand_high(141839.433982)

#Create a function for low demand at the current price
demand_high2<- function(x){alpha_high+beta_high*x - 5}

#Find the area under the curve via integrating
demandint_high <- integrate(demand_high2, 0, 141839.433982)
demandint_high #1304316

#Double check the area using shapes
high_height <- 23.39144 - 5
high_base <- 141839.433982
high_cs <- (1/2)*high_base*high_height #1304315.7199

#High demand benefit = 1304315.7199

```


Question 3: Gas tax

a) New quantity of gas
```{r}
# make a new supply cnurve with the 0.5 tax
supply_curve_tax<- function(x){supply*x+0.5}

# graph the new supply to see what it looks like

graph_tax<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_curve_tax, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))

graph_tax

# looks correct (supply shifted up by 0.5)

# find the new equilibrium with the new supply curve

q_opt<- uniroot(function(x) demand_agg(x) - supply_curve_tax(x), range(x))$root
q_opt

# New equilibrium qunatity = 261333

p_opt<- supply_curve_tax(q_opt)
p_opt

# new equilibrium price = 5.3899

# Q_tax = 261333
# P_tax = 5.3899


```

b) New gas price
```{r}
p_opt<- supply_curve_tax(q_opt)
p_opt

# new equilibrium price = 5.3899
```


c) New high consumer benefit
```{r}
#5.3899 = -0.0001296638(Q) + 23.39144
# Q = 138836.495

demand_high3<- function(x){alpha_high+beta_high*x - 5.3899}

demandint_high_tax <- integrate(demand_high3, 0, 138836.495)
demandint_high_tax

# geometric

demand_high(0)
# y - int = 23.39144

demand_high_base_tax<- 23.39144 - 5.3899
demand_high_height_tax<- 138836.495
demand_high_cs_tax<- (1/2)*demand_high_height_tax*demand_high_base_tax
demand_high_cs_tax

# CS for high demand with the new tax = 1249635

```

d) New low consumer benefit
```{r}
#5.3899 = -0.0001355174(Q) + 21.99085
# Q = 119,179.846

# geometric 

demand_low(0)
# y - int = 21.99085

demand_low_base_tax<- 21.99085 - 5.3899
demand_low_height_tax<- 119179.846
demand_low_cs_tax<- (1/2)*demand_low_base_tax*demand_low_height_tax
demand_low_cs_tax

# CS for low demand users with the new tax = 989249.3

```

e) New producer benefit
```{r}
# geometric 

supply_curve_tax(0)
# y - int = 0.5

supply_tax_base<- 5.3899-0.5
supply_tax_height<- 261333
ps_tax<- (1/2)*supply_tax_base*supply_tax_height
ps_tax

# PS under the new tax = 638946.1

```

f) Total envr damage
```{r}

mec_base_tax<- 261333
mec_height_tax<- 2
tec_tax<- mec_height_tax*mec_base_tax
tec_tax

# Total environmental cost under the new tax = 522666

```

g) Total tax revenue 
```{r}
# the length of the tax revenue rectanlge will just be the optinaml Q with the tax
# the height of the taxx revenue will be the difference between the optimal price before and after the tax
tax_rev_base<- 261333
tax_rev_height<- 5.3899 - 5
tax_rev<- tax_rev_height*tax_rev_base
tax_rev

# the revenue from the tax =  101893.7

```

4. Now, assume that all revenue from a tax will be used for infrastructure repairs,
and that the benefit of this is proportional to the amount you drive (i.e. to the gas
you consume). Also assume that “Low” income consumers bear all
environmental costs. For a range of gas taxes (ranging from $0 - $5.00/gal),
calculate the effects of the tax on:


```{r}
# set up the problem

tax_amount<- seq(0.25,0.5, by = 0.25)
# add this to the supply curve
supply_tax_new<- function(x){tax_amount + supply_curve*x}
# set price at the initial equilibrium of $5
tax_seq<- 5

# quantity with the tax:


# quantity with the tax
 ############################
# setting tax at 2
##########################
supply_tax_2<- function(x){supply*x + 2}

# graph for the new 2$ tax to visualize
graph_tax_2<- ggplot(data.frame(x = c(70000:500000)), aes(x=x))+
  stat_function(fun = demand_low, color = "darkblue")+
  stat_function(fun = demand_high, color = "darkorange")+
  stat_function(fun = demand_agg, color = "darkgreen")+
  stat_function(fun = supply_tax_2, color = "pink")+
  stat_function(fun = mec, color = "darkred")+
  xlab("\nQuantity")+
  ylab("Price\n")+
  ggtitle("Intial Gas Market \n")+
  scale_y_continuous(limits = c(0,25), expand = c(0,0))+
  scale_x_continuous(limits = c(0,500000), expand = c(0,0))
graph_tax_2
```

a: Overall welfare of “High” income consumers
```{r}
# consumer surplus for high consumers
tax_base_cs_high_2<- uniroot(function(x) demand_high(x) - supply_tax_2(x), range(x))$root
tax_height_cs_high_2<- alpha_high - demand_high(tax_base_cs_high_2)

tax_cs_area_high_2<- 0.5*tax_base_cs_high_2*tax_height_cs_high_2
tax_cs_area_high_2
# consumer surplus for high users for new tax = 1347555

```

b: Overall welfare of “Low” income consumers

```{r}
tax_base_cs_low_2<- uniroot(function(x) demand_low(x) - supply_tax_2(x), range(x))$root

tax_height_cs_low_2<- alpha_low - demand_low(tax_base_cs_high_2)

tax_cs_area_low_2<- 0.5*tax_base_cs_low_2*tax_height_cs_low_2
tax_cs_area_low_2
# consumer surplus for low demand group after tax = 1266222


```

c: Gas producers

```{r}
ps_low_tax_2<- 0.5*tax_base_cs_low_2*demand_low(tax_base_cs_low_2)
ps_low_tax_2
# producer surplus under the new 2 dollar tax for the low demand =  286801.9

ps_high_tax_2<- 0.5*tax_base_cs_high_2*demand_high(tax_base_cs_high_2)
ps_high_tax_2
# producer surplus under the new 2 dollar tax for the high demand = 338632.8

# aggregate demand producer surplus under the tax
ps_agg_tax_2<- ps_low_tax_2 + ps_high_tax_2
```

6. Using these numbers, which has a bigger impact on environmental quality, a
$2.00/gal. gas tax or the invention of electric cars? Provide your evidence.

```{r}







```






















